```{r setup}
library(tidyverse)
library(readxl)
library(kableExtra)
```
# Program Requirements 

Basic requirements to complete the PhD include (1) Required and elective coursework; (2) Passing a comprehensive exam; (3) Developing and defending an original thesis proposal; (4) Writing, submitting, and defending the thesis research.

## Coursework
PhD students are required to complete 25 course credits, including 16 required credits and 9 elective credits. 

### Required courses
The required coursework is typically completed during the first 4 terms and consists of the following courses:

- EPIB 701 PhD Comprehensive Examination*
- EPIB 702 PhD Proposal*
- EPIB 703 Principles of Study Design (2 Credits)
- EPIB 704 Doctoral Level Epidemiologic Methods 1 (4 Credits)
- EPIB 705 Doctoral Level Epidemiologic Methods 2 (4 Credits)
- EPIB 706 Doctoral Seminar in Epidemiology (3 Credits)
- EPIB 707 Research Design in the Health Sciences (3 Credits)

*Note that EPIB 701 and 702 are not didactic courses but are required milestones for advancing toward degree completion and require registration in the appropriate term. Students register for EPIB 702 in both Fall and Winter terms of their second year.

### Elective courses
Students are also required to take 9 credits of elective coursework, at the 500 level or higher, with a minimum of 3 credits in Biostatistics and 6 credits in an epidemiologic and/or substantive topic (normally related to the thesis topic). Elective courses must be chosen in consultation with the student's supervisor and/or the degree program’s director or adviser.

These courses can be chosen from the Department’s current offer of more than 40 courses in EBOH as well as from other McGill Departments. To assist you in your course selections see the Ph.D. Epidemiology Electives Guidelines page. Below you can find a list of current EBOH courses commonly taken as electives. However, courses from other departments or faculties may be possible, depending on the thesis subject matter and subject to the approval of your supervisor(s) and the Program Director. 

```{r elect, echo=FALSE}
d <- read_xlsx("docs/PhD-electives-list.xlsx", range = cell_cols("A:C"))
d1 <- drop_na(d)
kbl(d1, caption = "EBOH Electives as of Fall 2021") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Directed Reading Courses

**Directed Reading courses complement offerings in the department or elsewhere at McGill or other universities. They are NOT substitutes for existing courses but are, rather, ways for students in the programs to enrich their education in an organized way on topics not otherwise covered or not covered sufficiently (in depth or breadth) in existing courses.** 

Students enrolled in the department may take Directed Reading courses for credit towards a degree under the rubric of the Special Topics offerings. These courses may be for 1, 2, or 3 credits. Directed Reading courses should conform to the usual semester format unless the specific circumstances of the course require flexibility. However, students are expected to complete such a course within no more than any six month period.

While there is much flexibility in what constitutes such a course, certain requirements must be met before work can begin.

1. Students must themselves propose a supervising faculty member with whom to work.

2. With the faculty supervisor, students must prepare an adequate project proposal commensurate with the number of credits sought that includes:

a. The rationale for doing this work as a Directed Reading course and for the number of credits sought. As well, this statement should indicate how it relates to, but is separate from, thesis work when the student is in a thesis program.

b. An outline of the work to be done and the final product/output to be submitted. If a Reading Course is being proposed, a preliminary bibliography and a planned reading schedule should be included.

c. A timetable, with appropriate milestones to assess a student’s progress and the measures to be used to evaluate the work (e.g., number of written assignments and their length). A student’s faculty supervisor will be responsible for this evaluation as is the case for “regular” courses.

d. A timetable indicating when the student and faculty supervisor will meet.

Students will be expected to submit for approval in advance material that provides the objectives and methods to be used for the IS/DR work.

3. The project proposal, signed by both the student and the supervisor, should be submitted to the Student Affairs Office a minimum of one month prior to the start of the semester in which the course will take place. The director, along with one other person on the Program Committee who has accepted responsibility for curriculum matters, will review the proposal and determine if it is to be approved. Once approved internally, a copy will be sent to the Director of Graduate Studies as well as to the Department’s Graduate Studies Office, with a request that the latter obtain a Special Topics course number for the offering. A copy of the final approved version of the course content will be placed in the student’s file.

## Example curriculum
The timing and choices to fulfill the course requirements will likely be unique for each student. Below we provide one example of a possible curriculum over the course of the program.

- Year 1 
  - Fall: EPIB 703 Study Design; EPIB 704 Epi Methods I; Ethics Requirement: Tri-Council Policy for Ethical Conduct of Research online module (TCPS-2) (non-credit)  
  - Winter: EPIB 705 Epi Methods II; EPIB 706 Doctoral Seminar
  - Summer: EPIB 701 Comprehensive Exam (June)

- Year 2
  - Fall: EPIB 702 PhD proposal; EPIB 707 Research Design; BIOS elective (e.g., Advanced Generalized Linear Models, Causal Inference)
  - Winter: EPIB 702 PhD proposal; EPIB or substantive electives (e.g. Pharmacoepidemiology, Impact Evaluation,  Knowledge Synthesis)
  
## Concentration Coursework
As described previously, there are presently 3 options for PhD students that wish to pursue concentrated work in substantive areas related to either Global Health, Pharmacoepidemiology, or Population Dynamics. These options all require additional courses to be completed *in addition to* the required courses for all Epidemiology PhD students. 

### Global Health Option
This option will provide enhanced training in global health to graduate students registered in the Ph.D. in Epidemiology; Global Health degree program at McGill. Students will become familiar with topics of global health relevance and incorporate this into their core coursework and thesis research. The thesis must be relevant to global health and approved by the Global Health Coordinating Committee. Contextualizing the core training students receive in epidemiology and in their respective substantive discipline within the global health research domain will enhance their academic experience. Graduates of this option will be prepared to pursue further training in global health or to undertake a variety of career opportunities in global health in Canada or internationally.

Program Requirements

Thesis

A thesis for the doctoral degree must constitute original scholarship and must be a distinct contribution to knowledge. It must show familiarity with previous work in the field and must demonstrate ability to plan and carry out research, organize results, and defend the approach and conclusions in a scholarly manner. The research presented must meet current standards of the discipline; as well, the thesis must clearly demonstrate how the research advances knowledge in the field. Finally, the thesis must be written in compliance with norms for academic and scholarly expression and for publication in the public domain. The thesis must be relevant to global health and approved by the Global Health Coordinating Committee.

Required Courses (22 credits)

EPIB 681 Global Health: Epid. Research 3 Credits
EPIB 701 Ph.D.Comprehensive Examination
EPIB 702 Ph.D. Proposal
EPIB 703 Principles of Study Design 2 Credits
EPIB 704 Doctoral Level Epib Methods 1 4 Credits
EPIB 705 Doctoral Level Epid Methods 2 4 Credits
EPIB 706 Doctoral Seminar:Epidemiology 3 Credits
EPIB 707 Res Design in Health Sci 3 Credits
PPHS 511 Fundamentals of Global Health 3 Credits
Complementary Courses (9 credits)

6 credits of coursework at the 500 level or higher, with a minimum of 3 credits in biostatistics, and 3 credits in epidemiology. Courses must be chosen in consultation with the student's supervisor and/or the degree program's director or adviser.

3 credits of coursework at the 500 level or higher from this list, or any other course approved by the Global Health Option Committee that have not been taken to satisfy other program requirements.

GEOG 503 Advanced Topics in Health Geog 3 Credits
NUTR 501 Nutrition in Dev Countries 3 Credits
PPHS 525 HlthCare Systems in Comp Persp 3 Credits
PPHS 527 Econ for Hlth Serv Res&Policy 3 Credits
PPHS 529 Global Env Hlth&Burden/Disease 3 Credits
SOCI 513 Soc Aspects HIV/AIDS in Africa 3 Credits
SOCI 519 Gender and Globalization 3 Credits
SOCI 545 Sociology of Population 3 Credits
Comprehensive Exam (EPIB 701)

Students will normally take the Comprehensive Exam (EPIB 701) within 12 to 24 months of entry into the Ph.D. degree program. The Comprehensive Exam is held once a year in June. The exam is intended to test students’ ability to synthesize and integrate epidemiological knowledge. For details on the Comprehensive Exam, see the course outline for EPIB 701. The exam is graded “Pass” or “Fail”. Students with a “Fail” must repeat the exam the following year.

Protocol Defense (EPIB 702)

The comprehensive exam and the above required courses are usually completed before submitting and defending the thesis research protocol in EPIB 702. For details on the Protocol Defense, see the course outline for EPIB 702. The exam is graded “Pass” or “Provisional Pass” or “Fail”.

Students with a “Provisional Pass” (some deficiencies noted but not enough to stop progress on the thesis) are recommended to undertake (in consultation with their supervisor), specific remedial steps to address the areas of weakness identified in the exam. These could include additional courses, essays, assignments, and short courses. After satisfactory completion of these remedial steps the student will be considered to have a “Pass” on the Comprehensive Exam/Protocol Defense. Students are permitted to continue with their normal progression through the program.

Students with a “Fail” must repeat the exam the following year.

Thesis Research

Thesis research is normally actively undertaken following the comprehensive Exam. It is expected that students will complete their degree within 48 to 60 months of entry into the Ph.D. degree program.

GH Advisor:
Dr. M. Pai


### Pharmacoepidemiology Additional Reqired Courses
This program provides in-depth training for graduate students on pharmacoepidemiologic methods and the application of these methods to study the population effects (benefits and harm) of pharmaceutical products. Students will acquire the skills to become independent investigators and conduct original research in pharmacoepidemiology. Career opportunities for graduates are multiple and include work in industry, government, or academia. Students will be required to participate in the Pharmacoepidemiology Journal Club. Research topics must be related to pharmacoepidemiology and approved by the program coordinating committee.

Program Requirements

Thesis

A thesis for the doctoral degree must constitute original scholarship and must be a distinct contribution to knowledge. It must show familiarity with previous work in the field and must demonstrate ability to plan and carry out research, organize results, and defend the approach and conclusions in a scholarly manner. The research presented must meet current standards of the discipline; as well, the thesis must clearly demonstrate how the research advances knowledge in the field. Finally, the thesis must be written in compliance with norms for academic and scholarly expression and for publication in the public domain.

Required Courses (25 credits)

EPIB 639 Pharmacoepidemiologic Methods 4 Credits
EPIB 654 Pharmacoepidemiology 4 2 Credits
EPIB 661 Pharmacoepidemiology 3 2 Credits
EPIB 662 Pharma Basis of Pharmacoepidem 1 Credits
EPIB 701 Ph.D.Comprehensive Examination
EPIB 702 Ph.D. Proposal
EPIB 703 Principles of Study Design 2 Credits
EPIB 704 Doctoral Level Epib Methods 1 4 Credits
EPIB 705 Doctoral Level Epid Methods 2 4 Credits
EPIB 706 Doctoral Seminar:Epidemiology 3 Credits
EPIB 707 Res Design in Health Sci 3 Credits
Complementary Courses (3 credits)

3 credits of coursework in biostatistics at the 500 level or higher. Courses must be chosen in consultation with the student’s supervisor and/or the degree program’s director or adviser.

These courses can be chosen from the Department’s current offer of more than 40 courses in epidemiology, biostatistics and occupational health as well as from other McGill Departments. To assist you in your course selections see the Ph.D. Epidemiology Electives Guidelines page.

Comprehensive Exam (EPIB 701)

Students will normally take the Comprehensive Exam (EPIB 701) within 12 to 24 months of entry into the Ph.D. degree program. The Comprehensive Exam is held once a year in June. The exam is intended to test students’ ability to synthesize and integrate epidemiological knowledge. For details on the Comprehensive Exam, see the course outline for EPIB 701. The exam is graded “Pass” or “Fail”. Students with a “Fail” must repeat the exam the following year.

Protocol Defense (EPIB 702)

The comprehensive exam and the above required courses are usually completed before submitting and defending the thesis research protocol in EPIB 702. For details on the Protocol Defense, see the course outline for EPIB 702. The exam is graded “Pass” or “Provisional Pass” or “Fail”.

Students with a “Provisional Pass” (some deficiencies noted but not enough to stop progress on the thesis) are recommended to undertake (in consultation with their supervisor), specific remedial steps to address the areas of weakness identified in the exam. These could include additional courses, essays, assignments, and short courses. After satisfactory completion of these remedial steps the student will be considered to have a “Pass” on the Comprehensive Exam/Protocol Defense. Students are permitted to continue with their normal progression through the program.

Students with a “Fail” must repeat the exam the following year.

Thesis Research

Thesis research is normally actively undertaken following the comprehensive Exam. It is expected that students will complete their degree within 48 to 60 months of entry into the Ph.D. degree program.

PE Advisor:
Dr. R. Platt

### Population Dynamics Additional Required Courses
The Population Dynamics Option (PDO) is a cross-disciplinary, cross-faculty graduate program offered by the Centre on Population Dynamics (CPD) as an option within existing master’s and doctoral programs in the Departments of Sociology, Economics, and Epidemiology, Biostatistics and Occupational Health (EBOH) at McGill University. Students who have been admitted through their home department or faculty may apply for admission to the option. The option is coordinated by the CPD, in partnership with participating academic units.

Thus, in addition to the rigorous training provided in the Department of EBOH, graduate students who choose this option become Centre on Population Dynamics (CPD) student trainees. This affiliation notably offers opportunities for interdisciplinary research and supervision. The option also provides a forum whereby graduate students bring their disciplinary perspectives together and enrich each other's learning through structured courses, a weekly seminar series, and informal discussions and networking.

With interdisciplinary research being increasingly important to understanding complex social and biological processes, CPD student trainees benefit from both a strong disciplinary foundation from their departmental affiliations, as well as from the sharing of knowledge across disciplinary boundaries through CPD activities.

Program Requirements

Thesis

A thesis for the doctoral degree must constitute original scholarship and must be a distinct contribution to knowledge. It must show familiarity with previous work in the field and must demonstrate ability to plan and carry out research, organize results, and defend the approach and conclusions in a scholarly manner. The research presented must meet current standards of the discipline; as well, the thesis must clearly demonstrate how the research advances knowledge in the field. Finally, the thesis must be written in compliance with norms for academic and scholarly expression and for publication in the public domain.

Required Courses (22 credits)

EPIB 701 Ph.D.Comprehensive Examination
EPIB 702 Ph.D. Proposal
EPIB 703 Principles of Study Design 2 Credits
EPIB 704 Doctoral Level Epib Methods 1 4 Credits
EPIB 705 Doctoral Level Epid Methods 2 4 Credits
EPIB 706 Doctoral Seminar:Epidemiology 3 Credits
EPIB 707 Res Design in Health Sci 3 Credits
SOCI 545 Sociology of Population 3 Credits
SOCI 626 Demographic Methods 3 Credits
Complementary Courses (9 credits)

9 credits of coursework, at the 500 level or higher, with a minimum of 3 credits in biostatistics, 3 credits in epidemiology, and 3 credits from courses approved for the Population Dynamics Option from the list below:

ECON 622 Public Finance 3 Credits
ECON 634 Economic Development 3 3 Credits
ECON 641 Labour Economics 3 Credits
ECON 734 Economic Development 4 3 Credits
ECON 741 Advanced Labour Economics 3 Credits
ECON 742 Empirical Microeconomics 3 Credits
ECON 744 Health Economics 3 Credits
EPIB 648 Methods in Social Epidemiology 3 Credits
EPIB 681 Global Health: Epid. Research 3 Credits
PPHS 525 HlthCare Systems in Comp Persp 3 Credits
PPHS 528 Economic Eval of Hlth Programs 3 Credits
PPHS 529 Global Env Hlth&Burden/Disease 3 Credits
PPHS 615 Intro:Infectious Disease Epid 3 Credits
SOCI 502 Sociology of Fertility 3 Credits
SOCI 512 Ethnicity & Public Policy 3 Credits
SOCI 513 Soc Aspects HIV/AIDS in Africa 3 Credits
SOCI 520 Migration and Immigrant Groups 3 Credits
SOCI 525 HlthCare Systems in Comp Persp 3 Credits
SOCI 535 Sociology of the Family 3 Credits
SOCI 588 Biosociology/Biodemography 3 Credits
Courses must be chosen in consultation with the student's supervisor and/or the degree program's director or adviser.

Comprehensive Exam (EPIB 701)

Students will normally take the Comprehensive Exam (EPIB 701) within 12 to 24 months of entry into the Ph.D. degree program. The Comprehensive Exam is held once a year in June. The exam is intended to test students’ ability to synthesize and integrate epidemiological knowledge. For details on the Comprehensive Exam, see the course outline for EPIB 701. The exam is graded “Pass” or “Fail”. Students with a “Fail” must repeat the exam the following year.

Protocol Defense (EPIB 702)

The comprehensive exam and the above required courses are usually completed before submitting and defending the thesis research protocol in EPIB 702. For details on the Protocol Defense, see the course outline for EPIB 702. The exam is graded “Pass” or “Provisional Pass” or “Fail”.

Students with a “Provisional Pass” (some deficiencies noted but not enough to stop progress on the thesis) are recommended to undertake (in consultation with their supervisor), specific remedial steps to address the areas of weakness identified in the exam. These could include additional courses, essays, assignments, and short courses. After satisfactory completion of these remedial steps the student will be considered to have a “Pass” on the Comprehensive Exam/Protocol Defense. Students are permitted to continue with their normal progression through the program.

Students with a “Fail” must repeat the exam the following year.

Thesis Research

Thesis research is normally actively undertaken following the comprehensive Exam. It is expected that students will complete their degree within 48 to 60 months of entry into the Ph.D. degree program.

PDO Advisor:
Dr. A. Quesnel-Vallée